Deriving Multi-Headed Planar Dependency Parses from Link Grammar Parses

### 2014-12-07 doc/slides/hong+eisner.TLT13.slides.tex
comment
added more slides regarding the ILP, Experiments, and Conclusions section. ILP section is incomplete. Motivation section also needs work.

\section{Non Projectivity}
\frame{\frametitle{We do not address: Non-Projectivity}
\begin{itemize}
  \item Projectivity assumptions allow us to have efficient parsing algorithms that utilize dynamic programming
  \item English is mostly projective
\end{itemize}
}

@@

\frame{\frametitle{Link Grammars}
  \begin{enumerate}
    \item Grammar-based formalism for projective dependency parsing with undirected links
  \end{enumerate}
}

\frame{\frametitle{Link Grammars}
\begin{figure}
  \begin{dependency}[edge style={-}]
	\begin{deptext}
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	\depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	\depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	\depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	\depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	\depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	\depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	\deproot[edge above, thick, edge style = {red}]{2}{W}
	\deproot[edge above, thick, edge style = {red}]{7}{WV}
	\deproot[edge above, thick, edge style = {red}]{10}{X}
	\depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	\depedge[edge above, thick, edge style = {red}]{6}{5}{E}
  \end{dependency}
  \caption*{Link Parse of a sentence from Penn Tree Bank}
\end{figure}

@@

\frame{\frametitle{Link Grammars}
\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	\depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	\depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	\depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	\depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	\depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	\depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	\deproot[edge above, thick, edge style = {red}]{2}{W}
	\deproot[edge above, thick, edge style = {red}]{7}{WV}
	\deproot[edge above, thick, edge style = {red}]{10}{X}
	\depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	\depedge[edge above, thick, edge style = {red}]{6}{5}{E}
  \end{dependency}

@@

  \caption*{Compare resulting dependency parse with Penn Treebank annotations}
\end{figure}

@@

\section{ILP}
\frame{\frametitle{Integer Linear Programming}

\NoteJH{Will the audience know about ILP?}

@@

\section{ILP Model}
\frame{\frametitle{Integer Linear Programming Model}

\NoteJH{Take from the paper.}

@@

\section{Experiments and Results}
\frame{\frametitle{Experiments and Results}

@@

\section{Conclusions}
\frame{\frametitle{Conclusions}

>>>>

\section{Non Projectivity}
\frame{\frametitle{We do not address: Non-Projectivity}
\begin{itemize}
  \item Projectivity assumptions allow us to have efficient parsing algorithms that utilize dynamic programming
  \item English is mostly projective
\end{itemize}

\NoteJH{Should I mention non-projectivity? This slide is mostly here because of a reviewer comment.}
}

@@

\frame{\frametitle{Link Grammars}
  \begin{itemize}
    \item Grammar-based formalism for projective dependency parsing with undirected links
    \item We set out to discover if the undirected links are actually directed relationships.
  \end{itemize}
}

\frame{\frametitle{Link Grammars: Example Parse}
\begin{figure}
  \begin{dependency}[edge style={-}]
	\begin{deptext}
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	\depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	\depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	\depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	\depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	\depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	\depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	\deproot[edge above, thick, edge style = {red}]{2}{W}
	\deproot[edge above, thick, edge style = {red}]{7}{WV}
	\deproot[edge above, thick, edge style = {red}]{10}{X}
	\depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	\depedge[edge above, thick, edge style = {red}]{6}{5}{E}
  \end{dependency}
  \caption*{Link Parse of a sentence from Penn Tree Bank}
\end{figure}

@@

\frame{\frametitle{Link Grammars: Converting into a Directed Acyclic Graph}
\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	\depedge[edge above, thick, edge style = {red}]{2}{3}{S}
	\depedge[edge above, thick, edge style = {red}]{3}{6}{I}
	\depedge[edge above, thick, edge style = {red}]{2}{1}{D}
	\depedge[edge above, thick, edge style = {red}]{7}{8}{MV}
	\depedge[edge above, thick, edge style = {red}]{6}{7}{P}
	\depedge[edge above, thick, edge style = {red}]{8}{9}{J}
	\deproot[edge above, thick, edge style = {red}]{2}{W}
	\deproot[edge above, thick, edge style = {red}]{7}{WV}
	\deproot[edge above, thick, edge style = {red}]{10}{X}
	\depedge[edge above, thick, edge style = {red}]{6}{4}{E}
	\depedge[edge above, thick, edge style = {red}]{6}{5}{E}
  \end{dependency}

@@

  \caption*{Compare resulting dependency parse with CoNLL 2007 shared task}
\end{figure}

@@

%\section{ILP}
%\frame{\frametitle{Integer Linear Programming}

%\NoteJH{Will the audience know about ILP?}

%}

\section{ILP Model}
\frame{\frametitle{Integer Linear Programming Model: Variables}

For each sentence, for each edge $i,j$ where $i < j$, we introduce variables $x_{ij}$ and $x_{ji}$.

\NoteJH{TODO: Incomplete slide. Need to describe things like depth of the variable, label}
}

\frame{\frametitle{Integer Linear Programming Model: Encoded Constraints}

\begin{itemize}
  \item Connectedness
  \item Acyclicity
  \item Consistency of Directionalized Links
\end{itemize}

\NoteJH{Take from the paper.}
}

\subsection{Connectedness, Acyclicity}
\frame{\frametitle{Connectedness, Acyclicity}
\begin{itemize}
  \item[] Connectedness
    \begin{align}
      \sum_u x_{uv} & \geq 1
    \end{align}

  \item[] Acyclic
    \begin{align}
      (\forall_u)\; d_v + (1 + n_v) \cdot (1 - x_{uv}) & \geq 1+d_u
    \end{align}
\end{itemize}

@@

\frame{\frametitle{Consistence of Directionalized Links}
\begin{itemize}
  \item[] Consistency of Directionalized Links
    \begin{align}\label{direction+slack}
      x_{ij} &\leq r_L + s_{ij} &
      x_{ji} &\leq \ell_L + s_{ij}
    \end{align}

    \begin{align}\label{eqn:obj}
      \min \left( \sum_L r_L + \ell_L \right)
    \end{align}

    %$$x_{ij} \leq r_L + s_{ij} $$
    %$$x_{ji} \leq \ell_L + s_{ij}$$
    %$$\min \left( \sum_L r_L + \ell_L \right)$$

\end{itemize}
}

\frame{\frametitle{Consistence of Directionalized Links with Slack}
\begin{itemize}
  \item[] Consistency of Directionalized Links
    \begin{align}\label{direction+slack}
      x_{ij} &\leq r_L + s_{ij} &
      x_{ji} &\leq \ell_L + s_{ij}
    \end{align}

    \begin{align}\label{eqn:obj}
      \min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}
    \end{align}

    %$$x_{ij} \leq r_L + s_{ij} $$
    %$$x_{ji} \leq \ell_L + s_{ij}$$
    %$$\min \left( \sum_L r_L + \ell_L \right) \cdot \frac{N_L}{4} + \sum_{ij}s_{ij}$$

\end{itemize}
}

\section{Experiments and Results}
%\frame{\frametitle{Experiments and Results}
%\begin{itemize}
%  \item 18,577 English sentences with gold CoNLL. 
%  \item 18,577 unlabeled Russian sentences.
%\end{itemize}
%}

\frame{\frametitle{Data Sets}
\begin{itemize}
  \item[] 18,577 English sentences with 10,960 connected parses
  \item[] 18,577 unlabeled Russian sentences with 4,913 connected parses
\end{itemize}

@@

\frame{\frametitle{On the English data set}
\begin{itemize}
\item[] Link Data has 8\% additional edges over the CoNLL.
\item[] 52\% of links match CoNLL arcs
\item[] 57\% of CoNLL arcs match links
\end{itemize}

\NoteJH{I skipped the ``Stability of Results'' section}
}

\frame{\frametitle{ILP Results}
  \begin{itemize}
    \item[] Link types that allowed both directions:
      \begin{itemize}
        \item[] $\frac{7}{113} = 6.19\%$ 
      \end{itemize}

    \item[] Link tokens that required disallowed direction via slack:
      \begin{itemize}
        \item[] $\frac{4043}{195000} = 2.07\%$ 
      \end{itemize}
    \item[]
  \end{itemize}
}

\frame{\frametitle{ILP Results}
  \begin{itemize}
    \item[] Link types that allowed both directions:
      \begin{itemize}
        \item[] $\frac{7}{113} = 6.19\%$ 
      \end{itemize}

    \item[] Link tokens that required disallowed direction via slack:
      \begin{itemize}
        \item[] $\frac{4043}{195000} = 2.07\%$ 
      \end{itemize}
    \item[] The link labels mostly have a consistent direction.
  \end{itemize}
}

\frame{\frametitle{Link Results}
  \begin{itemize}
    \item Subject-Verb links are backwards.
  \end{itemize}

\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
	\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
	\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
	\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
	\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
	\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
	\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
	\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
	\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
	\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
	\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
	\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
	\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
	\deproot[edge below, edge style = {red, dotted}]{2}{W}
	\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
	\deproot[edge below, edge style = {red, dotted}]{10}{X}
	\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
	\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
  \end{dependency}
\end{figure}

\begin{itemize}
\item[] %This is probably due to an inconsistency of the Link Grammar, discovered by our method.
\end{itemize}

@@

\frame{\frametitle{Link Results}
  \begin{itemize}
    \item Subject-Verb links are backwards.
  \end{itemize}

\begin{figure}
  \begin{dependency}
	\begin{deptext}
	  {\scriptsize DT} \& {\scriptsize NN} \& {\scriptsize MD} \& {\scriptsize RB} \& {\scriptsize RB} \& {\scriptsize VB} \& {\scriptsize VB} \& {\scriptsize IN} \& {\scriptsize NN} \& {\scriptsize .} \\
	  the \& matter \& may \& never \& even \& be \& tried \& in \& court \& . \\
	  - \& n-u \& v \& e \& e \& v \& v-d \& r \& n-u \& - \\
	\end{deptext}
	%\deproot[edge above, edge style = {blue, dotted}]{3}{\small ROOT}
	\depedge[edge below, edge style = {red, ultra thick}]{2}{3}{S}
	\depedge[edge above, edge style = {blue, ultra thick}]{3}{2}{\small SBJ}
	%\depedge[edge above, edge style = {blue, dotted}]{3}{4}{\small ADV}
	%\depedge[edge above, edge style = {blue, dotted}]{3}{5}{\small ADV}
	%\depedge[edge below, edge style = {red, thick}]{3}{6}{I}
	%\depedge[edge above, edge style = {blue, thick}]{3}{6}{\small VC}
	%\depedge[edge above, edge style = {blue, dotted}, edge unit distance =1.5ex]{3}{10}{\small P}
	%\depedge[edge below, edge style = {red, thick}]{2}{1}{D}
	%\depedge[edge above, edge style = {blue, thick}]{2}{1}{\small NMOD}
	%\depedge[edge below, edge style = {red, thick}]{7}{8}{MV}
	%\depedge[edge above, edge style = {blue, thick}]{7}{8}{\small ADV}
	%\depedge[edge below, edge style = {orange, thick}]{6}{7}{P}
	%\depedge[edge above, edge style = {blue, thick}]{6}{7}{\small VC}
	%\depedge[edge below, edge style = {red, thick}]{8}{9}{J}
	%\depedge[edge above, edge style = {blue, thick}]{8}{9}{\small PMOD}
	%\deproot[edge below, edge style = {red, dotted}]{2}{W}
	%\deproot[edge below, edge style = {orange, ultra thick, dotted}]{7}{WV}
	%\deproot[edge below, edge style = {red, dotted}]{10}{X}
	%\depedge[edge below, edge style = {red, dotted}]{6}{4}{E}
	%\depedge[edge below, edge style = {red, dotted}]{6}{5}{E}
  \end{dependency}
\end{figure}

\begin{itemize}
\item This is due to an inconsistency of the Link Grammar, discovered by our method. \NoteJH{The reason why is in footnote 6 of page 10 in our paper. Reread after sleeping.}
\end{itemize}

\NoteJH{Should I include a section of the appendix table? Which rows should I include?}
}

\section{Conclusions}
\frame{\frametitle{Conclusions}
  \begin{itemize}
  \item Link Grammar parses can be oriented into connected DAGs
  \item New corpora for building multi-headed dependency parsers
  \item ILP can be used to build or annotate corpora
  \end{itemize}
}

\frame{\frametitle{}
  \begin{itemize}
  \item[] Questions?
  \end{itemize}
}
